\documentclass[11pt]{article}

\usepackage{amsfonts,amsmath,amssymb,latexsym,color,epsfig}
\setlength{\textheight}{22.5cm} \setlength{\textwidth}{6.7in}
\setlength{\topmargin}{0pt} \setlength{\evensidemargin}{1pt}
\setlength{\oddsidemargin}{1pt} \setlength{\headsep}{10pt}
\setlength{\parskip}{1mm} \setlength{\parindent}{0mm}
\renewcommand{\baselinestretch}{1.03}

\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newenvironment{proof}[1]
      {\medskip\noindent{\bf Proof #1}\hspace{1mm}}
      {\hfill$\Box$\medskip}


\title{\vspace{-0.7cm}Ma 1A - Problem Set 1}
\date{}

\begin{document}

\maketitle

\begin{enumerate}

\item
Prove that $x + \frac 1x \geq 2$ for all $x > 0$ and use this result to show that $\frac{a+b}{2} \geq \sqrt{ab}$ for all $a, b > 0$.

\bigskip 
\textbf{Proof: } First show that $x + \frac 1x \geq 2$ for all $x > 0$. Start by multiplying both sides of the inequality by $x$, which is ok because we only care about $x\geq 0$ and don't have to worry about sign switching or multiplying by a negative number: 
\[x^2 + 1 \geq 2x\]
Subtract $2x$ from both sides: 
\[x^2 -2x + 1 \geq 0 \]
Now let's find the minimum value of the left side of the inequality. Take the derivative and set it equal to zero: 
\[\tfrac{d}{dx} (x^2-2x+1 )= 2x -2 = 0 \]
\[2x = 2 \]
\[x = 1 \]
Thus we know that the minimum value of $x^2 - 2x + 1$ occurs at $x = 1$. Plug in $x = 1$ to the inequality to find the minimum value: 
\[(1)^2 - 2(1) +1 \geq 0 \]
\[ 1 - 2 + 1 \geq 0\]
\[0 \geq 0 \]
Which is true. Thus, because we know that the left side of the inequality is greater than or equal to $0$ at its minumum, then we can infer that the left side of the inequality is greater than or equal to $0$ at every value of $x$ and conclude that the inequality is true. 

\medskip

Now move on to the statement $\frac{a+b}{2} \geq \sqrt{ab}$ for all $a, b > 0$. Square both sides of the inequality: 
\[\frac{(a+b)^2}{4} \geq ab \]
Expand the top of the left side of the inequality: 
\[\frac{a^2 + 2ab + b^2}{4} \geq ab \]
Multiply both sides by 4: 
\[a^2 + 2ab + b^2 \geq 4ab\]
Subtract $2ab$ from both sides:
\[ a^2 + b^2 \geq 2ab\]
Divide both sides by $2ab$ ($a$ and $b$ are given to be always positive so no inequality sign switching takes place):
\[\frac{a^2 + b^2}{ab} \geq 2 \]
Split the fraction apart and simplify:
\[\frac{a^2}{ab} + \frac{b^2}{ab} \geq 2\]
\[\frac{a}{b}+ \frac{b}{a} \geq 2 \]
Plug in $x = \frac{a}{b}$:
\[x + \frac{1}{x} \geq 2 \] 
Which we know is true from the first part of this question. Thus, the statement $\frac{a+b}{2} \geq \sqrt{ab}$ for all $a, b > 0$ is true. 

\bigskip
\rightline{$\Box$}

\item
Prove that $\sqrt{10}$ is irrational and deduce that $\sqrt{2} + \sqrt{5}$ is also irrational.

\bigskip 
\textbf{Proof: } First prove that $\sqrt{10}$ is irrational. This will be a proof by contradiction. Assume to the contrary that $\sqrt{10}$ is rational and can be represented by an irreducible $\frac{p}{q}$ for $p,q \in \mathbb{Z}$, where $p$ and $q$ share no common factors. Then, 
\[\sqrt{10} = \frac{p}{q} \]
\[10 = \frac{p^2}{q^2}\]
\[p^2 = 10q^2 \]
 Note that if $p^2$ is a multiple of $10$, then $p$ is a multiple of $10$. This is because the prime factorization of $10$ is $2$ and $5$, both of which are prime numbers and not squares. Because $p$ is a multiple of 10, rewrite it as $10p_1, p_1 \in \mathbb{Z}$: 
\[100p_1 ^2 = 10 q^2\]
\[10p_1^2 = q^2   \]
Note that if $q^2$ is a multiple of $10$, then $q$ is also a multiple of $10$ and discussed above. Thus, both $p$ and $q$ share the common factor of $10$.
However, this is a contradiction, because we initially assumed that $p$ and $q$ shared no common factor. Thus, $\sqrt{10}$ cannot be rational and must instead be irrational. 

Now deduce that $\sqrt{2} + \sqrt{5}$ is also irrational. Assume to the contrary that $\sqrt{2} + \sqrt{5}$ is rational. Then, 
\[(\sqrt{2} + \sqrt{5})^2  = 2 + 2(\sqrt{2})(\sqrt{5}) + 5 = \]
\[ = 7 + 2\sqrt{10}\]
Let this supposedly rational number be represented by $q \in \mathbb{Q}$:
\[q = 7 + 2\sqrt{10}\]
Rearrange:
\[\frac{q-7}{2} = \sqrt{10}\]
However, $\frac{q-7}{2}$ should be rational because every number in it is rational, which creates a contradiction because we just showed that $\sqrt{10}$ is irrational. Thus, $\sqrt{2} + \sqrt{5}$ is also irrational.

\bigskip
\rightline{$\Box$}
\item
Prove, by induction or otherwise, that $\sum_{k=1}^n k^3 = \left(\sum_{k=1}^n k\right)^2$.

\bigskip\textbf{Proof: }

First examine the base case $n = 1$. Then, 
\[ \sum_{k=1}^1 k^3 = \left(\sum_{k=1}^1 k\right)^2 \]
\[ 1^3 = 1^2\]
\[ 1= 1\]

Now assume that $\sum_{k=1}^n k^3 = \left(\sum_{k=1}^n k\right)^2$ and show that the $n+1$ case holds true: 
\[ \sum_{k=1}^{n+1} k^3 = \left(\sum_{k=1}^{n+1} k\right)^2 \]
Seperate the last term of each series:
\[ \sum_{k=1}^{n} k^3 + (n+1)^3 = \left(\sum_{k=1}^n k + (n+1)\right)^2\]
Apply the inductive hypothesis: 
\[ \left(\sum_{k=1}^n k\right)^2 + (n+1)^3 = \left(\sum_{k=1}^n k + (n+1)\right)^2\]
Do binomial expansion on the right side:
\[ \left(\sum_{k=1}^n k\right)^2 + (n+1)^3 = \left( \sum_{k=1}^n k\right)^2 + 2\left( \sum_{k=1}^{n} k\right)(n+1) + (n+1)^2\]
Subtract $\left(\sum_{k=1}^n k\right)^2$ from both sides:
\[ (n+1)^3 = 2\left( \sum_{k=1}^{n} k\right)(n+1) + (n+1)^2\]
Side note: We're going to follow in the footsteps of 10 year old Gauss and prove that $\sum_{k=1}^{n} k = \frac{n(n+1)}{2}$. First prove the base case $n=1$:

\[\sum_{k=1}^{1} k =  \frac{1(1+1)}{2} \Rightarrow 1 = 1\]
Now prove the inductive step. Assume that $\sum_{k=1}^{n} k = \frac{n(n+1)}{2}$ and prove that $\sum_{k=1}^{n+1} k = \frac{(n+1)(n+2)}{2}$: 
\[ \sum_{k=1}^{n+1} k = \frac{(n+1)(n+2)}{2} \]
\[ \sum_{k=1}^{n} k + (n+1) = \frac{(n+1)(n+2)}{2} \]
Apply the inductive hypothesis:
\[ \frac{n(n+1)}{2} + (n+1) = \frac{(n+1)(n+2)}{2} \]
\[ \frac{n^2 + n + 2n + 2}{2}= \frac{n^2 + 3n + 2}{2}\]
\[ \frac{n^2 + 3n + 2}{2} = \frac{n^2 + 3n + 2}{2}\]
And thus we have proven through induction that $\sum_{k=1}^{n} k = \frac{n(n+1)}{2}$. 

Ok now back to the original problem. Apply the $\frac{n(n+1)}{2}$ identity+ to our previous progress:
\[ (n+1)^3 = 2(\frac{n(n+1)}{2})(n+1) + (n+1)^2\]
Now for some algebra:
\[ n^3 + 3n^2 +3n + 1 = n(n+1)(n+1)+ n^2 + 2n + 1 \]
\[ n^3 + 3n^2 +3n + 1 = (n^2+n)(n+1) + n^2 + 2n + 1\]
\[ n^3 + 3n^2 +3n + 1 = n^3 + n^2 + n^2 + n + n^2 + 2n + 1\]
\[ n^3 + 3n^2 +3n + 1 = n^3 + 3n^2 + 3n + 1\] 
And therefore the inductive case is proven. Because the base case and the inductive case are proven, the statement $\sum_{k=1}^n k^3 = \left(\sum_{k=1}^n k\right)^2$ is proven for all $n \geq 1$.

\rightline{$\Box$}
\item
Each card in a pack has a number on one side and a letter on the other. Four cards are
placed on the table:
\[ \boxed{2} \; \boxed{3} \; \boxed{A} \; \boxed{B}\]
You are permitted to turn just two cards in order to test the following hypothesis: a card that
has an even number on one side has a vowel on the other. Which two cards should you turn and why? 

\bigskip\textbf{Answer: } You only care about the statement that if a card has an even number on one side, then it has a vowel on the other. You don't care about what the card with 3 on one side has, for instance, because it doesn't have an even number on one side. Even if it ended up having a vowel on the other side, it would not invalidate our claim because it is not an ``iff" claim. Same goes for the card with the $A$ on the back, because we are not evaluating the claim ``If there is a vowel on one side, the other side is even''. There are only two cards that the hypothesis concerns: the card that has the even number $2$ on one side and the card that has the letter $B$ on one side. We care about the $B$ card beacuse it allows us to test the contrapositive to the hypothesis (which is logically equivalent), ``If there is not a vowel on the card, then there is not an even number on the back". We can use our two card turns to check these two cards and confirm or reject our hypothesis. 

\item
Which of the following relations on $\mathbb{N}$ are reflexive, which are symmetric and which are transitive:
\begin{itemize}
\item[(i)] $a \mid b$ (`$a$ divides $b$'),

\bigskip
For $a \mid b$, then there exists an integer $k$ such that $ka = b$.

\medskip
$a = 1a$, so $a \mid a$, so $\mid$ is reflexive. 

\medskip 
However, $\mid$ is not symmetric. Counterexample: $4$ divides $8$ but $8$ does not divide $4$. 

\medskip $\mid$ is transitive. Proof: if $a \mid b$, then $b = ka, k\in \mathbb{N}$. If $b \mid c$, then $c = pb, p\in\mathbb{N}$. Then, $c = pb = pka$. Thus, $a \mid c$. 
\item[(ii)] $a \nmid b$ (`$a$ does not divide $b$'),

\bigskip $\nmid$ is not reflexive, because we have already shown above that $a \mid a$, so it cannot be the case that $a \nmid a$.

\medskip 
$\nmid$ is not symmetric. For instance, $8 \nmid 4$ but $4 \mid 8$. 

\medskip 
$\nmid $ is not transitive. For instance, $3 \nmid 5$ and $5\nmid 9$ but $3 \mid 9$. 


\item[(iii)] $a$ and $b$ have the same remainder after division by $2025$,

\bigskip
Having the same remainder is reflexive. $a\mod 2025 = a\mod 2025$.

\medskip
Having the same remainder after division by 2025 is symmetric. If $a \mod 2025 = b \mod 2025$, then $b \mod 2025 = a \mod 2025$.

\medskip
Having the same remainder after division by 2025 is transitive. If $a \mod 2025 = b \mod 2025$, and $b \mod 2025 = c \mod 2025$, then $a \mod 2025 = c \mod 2025$.
\item[(iv)] gcd$(a, b) > 2025$?

\bigskip
If this statement were reflexive, then $\gcd(a,a)$ would be greater than 2025. Counterexample: When $a = 2$, $\gcd(2,2) = 2 \leq 2025$. 

If this statement were symmetric, then when $\gcd(a,b) > 2025$, $\gcd(b,a) > 2025$. This is true because $\gcd(a,b) = \gcd(b,a)$. 

If this statement were transitive, then if $\gcd(a,b) > 2025 $ and $\gcd(b,c) > 2025$ then $\gcd(a,c) > 2025$. This is disproven by the counterexample of $a = 2027, b = 2029*2027, c = 2029$. $\gcd(a,b) = 2027 > 2025, \gcd(b,c) = 2029 > 2027, $but$ \gcd(a,c) = 1 \leq 2025$. Thus this statement is not transitive. Note that $2027$ and $2029$ are prime numbers.
\end{itemize}

\end{enumerate}


\end{document}